\section{Managed bibliography with ADS
keys}\label{managed-bibliography-with-ads-keys}

Forget about bibliography management with pure NASA ADS keys!

This drop-in Perl script uses citation keys from the
\href{https://ui.adsabs.harvard.edu/}{NASA astronomy data service (ADS)}
so that you can forget about your bibliography management. For citation
commands in your document that use ADS keys, such as
\texttt{\textbackslash{}cite\{1958ZA.....46..108B\}}, the script
automatically fetches and caches the corresponding bibliography entries
in a managed bibliography file, normally with extension
\texttt{.adskeys.bib}. It also keeps a generated \texttt{.map} sidecar
with cached ADS key checks and canonicalization results.

Citations with keys that are not present in the ADS database can be
added by providing one or more additional bibliography files, as in
\texttt{\textbackslash{}bibliography\{paper.adskeys,custom\}}.

The script is intended for integration with
\href{https://ctan.org/pkg/latexmk/}{Latexmk}, which automates the build
process for LaTeX documents.

\subsection{How to use}\label{how-to-use}

Copy \href{./managed-bibliography.pl}{\texttt{managed-bibliography.pl}}
into your document directory. Then copy the
\href{./latexmkrc}{\texttt{latexmkrc}} file as well, or merge its
content with your existing \texttt{latexmkrc} file.

Add the managed bibliography file to your
\texttt{\textbackslash{}bibliography\{...\}} command. By default, a
document \texttt{paper.tex} uses \texttt{paper.adskeys.bib}, so a
typical setup looks like
\texttt{\textbackslash{}bibliography\{paper.adskeys,custom\}}. If you
set \texttt{\$managed\_bib\_file} to another filename, use that basename
instead.

Provide an ADS \href{https://ui.adsabs.harvard.edu/help/api/}{API token}
through \texttt{ADS\_API\_TOKEN}, \texttt{ADS\_DEV\_KEY},
\texttt{\$HOME/.ads/token}, or \texttt{\$HOME/.ads/dev\_key}. Avoid
putting tokens directly in a committed \texttt{latexmkrc} unless you
really need to.

The script uses \texttt{HTTP::Tiny} and \texttt{JSON::PP}. Perl 5.14 or
later is a sensible baseline. Then build as usual with \texttt{latexmk}.
On the first run the managed bibliography file and its \texttt{.map}
sidecar are created automatically.

\subsection{How it works}\label{how-it-works}

During a build, the script reads the \texttt{.aux} file to discover
which bibliography files are in use and which citation keys were
requested. It compares those keys with the managed bibliography file and
any user-supplied \texttt{.bib} files already listed in
\texttt{\textbackslash{}bibliography\{...\}}.

Keys that are still unresolved are sent to ADS. The script first asks
ADS to canonicalize identifiers, then exports the matching BibTeX
entries in bulk, rewrites the exported entry keys back to the keys cited
in the document, and stores the results in the managed bibliography
file. The generated \texttt{.map} sidecar stores cached ADS key checks
and canonicalization results so that duplicate warnings and ADS-valid
user-bibliography keys can be reported again on later runs without
re-querying ADS.

If two cited keys resolve to the same ADS entry, or if a cited key
resolves to a canonical ADS key that already exists in another
bibliography file, the script reports that. The \texttt{latexmkrc}
setting \texttt{\$allow\_duplicates} controls whether those cases are
allowed with a warning or stop the build.

If ADS does not return an entry, the key is left unresolved and
LaTeX/BibTeX behaves as it normally would for a missing citation.

By default, \texttt{latexmk\ -C} also removes the managed bibliography
file and its \texttt{.map} sidecar. The next build then performs a fresh
ADS lookup, which can refresh stale cached entries.

\subsection{Test case}\label{test-case}

To build \href{./manbib.tex}{\texttt{manbib.tex}} with Latexmk, run:

\begin{verbatim}
latexmk -pdf -bibtex manbib.tex
\end{verbatim}

This creates \href{./manbib.pdf}{\texttt{manbib.pdf}} together with the
managed bibliography file
\href{./manbib.adskeys.bib}{\texttt{manbib.adskeys.bib}} and the sidecar
cache \href{./manbib.adskeys.map}{\texttt{manbib.adskeys.map}}. To clean
up the generated files, run:

\begin{verbatim}
latexmk -C manbib.tex
\end{verbatim}

\subsection{Known issues}\label{known-issues}

If ADS lookups abort immediately with a message about missing Perl HTTPS
support, \texttt{latexmk} is probably running under a Perl that does not
have the required SSL modules. This often happens when \texttt{perl} on
your \texttt{PATH} comes from Conda rather than the macOS system Perl.

The script uses \texttt{HTTP::Tiny} for ADS requests, and HTTPS support
requires the Perl modules \texttt{IO::Socket::SSL} and
\texttt{Net::SSLeay}. If those modules are missing, \texttt{manbib}
aborts ADS work for that run and reports the active Perl runtime.

If this happens, check \texttt{which\ perl} in the same shell that runs
\texttt{latexmk}. The important point is not which Perl you use, but
that the active Perl runtime must have HTTPS support available to
\texttt{HTTP::Tiny}. If a non-system Perl is being used, such as one
from Conda, install the required Perl SSL modules there or adjust your
environment so that the Perl used by \texttt{latexmk} already has HTTPS
support.
